\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-06-06 doc/paper.tex
doc/paper.tex: avoid redundancy
sharply ~ manifestly
divided ~ separate
addressing Dominik's suggestions on section 7

The large variety of supersymmetric models and potential
investigations makes it likely that the user wants to modify the
generated spectrum generator source code or reuse components in
further programs.  \fs offers two levels to influence the code: (i) On
the \mathematica model file level the model itself or GUT/weak scale
boundary conditions as well as input and output parameters can be
controlled (see \secref{sec:quick-start-alternative-models} and
\secref{sec:adapting-model-files} for examples).  (ii) In particular
\fs uses C++ object orientation features to modularize the source code
so that it is sharply divided into manifestly separate building blocks.
This modular architecture makes it easy
for the user to modify, reuse, replace or
extend the individual components (see \secref{sec:adapting-cpp-code}
for examples).  An important application of this concept are the
boundary conditions, for which the C++ level offers a wider range of
possibilities.  The boundary conditions solver provides a plugin
mechanism via a common \code{Constraint} interface, which allows a
user to exchange or add boundary conditions at any scale.
To realized this, all (derived) constraint objects are intentionally
kept outside the solver.  Despite being independent of one another,
they can fit together with the aid of class inheritance.  An
elaborate example of a tower of effective field theories and multiple
matching scales is presented in \secref{sec:tower construction}.
Alternatively, the modular structure makes it straightforward to take
\fs generated code for e.g.\ RGEs or self-energies and reuse it in an
existing code for some other purpose.  Conversely, it is also easy to
include code from elsewhere into the spectrum generator.  For an
example see \secref{sec:integrating-custom-built}.

@@

By definition, research is an endeavor to find something new.
Therefore, it can often be the case that
a spectrum generator right out of the box is not enough.
\fs attempts to offer a clean interface through which
one can exploit its facilities
while undergoing a minimal amount of frustration,
when he or she programs for a wide variety of studies.
In what follows,
this property shall be demonstrated by presenting a few use cases
at differing levels of complexity.

@@

DEFINITION[EWSB][DiracSpinors] = {
  Fu  -> {FUL, conj[FUR]},
  Fv  -> {FV , conj[FV] },
  Chi -> {L0 , conj[L0] },
  Cha -> {Lm , conj[Lp] },
  ...
};
\end{numlstlisting}
With respect to the MSSM file, newly added lines are
6, 12, 15, and 22.
Notice that the (left- and right-handed) neutrino mixing
in line 15 resembles the neutralino mixing in line 14.
Due to the Majorana mass term in the superpotential,
the six neutrino mass eigenstates are described in terms of
Majorana spinors like the neutralinos.

@@

WeylFermionAndIndermediate = {
  {v,   { Description -> "Right Neutrino Superfield" }},
  {FV,  { Description -> "Neutrino-Masseigenstate"}},
  {FvL, { Description -> "Left Neutrino"}},
  {FvR, { Description -> "Right Neutrino"}},
  ...
};
\end{numlstlisting}
In line 12, one finds PDG codes beginning with \code{99}.
Such numbers are available for a program author's private use
\cite{Beringer:1900zz}.
The new parameters in the superpotential
and the soft supersymmetry breaking sector
are to be described in \code{parameters.m}:
\begin{numlstlisting}
ParameterDefinitions = {
  {UV,    { Description -> "Neutrino-Mixing-Matrix"}},
  {Yv,    { Description -> "Neutrino-Yukawa-Coupling" }},
  {T[Yv], { Description -> "Trilinear-Neutrino-Coupling"}},
  {Mv,    { LaTeX -> "M_v", OutputName -> Mv, LesHouches -> Mv}},
  {B[Mv], { LaTeX -> "B_v", OutputName -> BMv, LesHouches -> BMv}},
  {mv2,   { Description -> "Softbreaking right Sneutrino Mass"}},
  ...
};
\end{numlstlisting}
For details on how to write model files for \sarah, consult its manual.

@@

There are problems which one cannot solve only by
editing \mathematica model files.
To unlock the full potential of \fs,
it is an advantage not to avoid programming at the C++ level.
For this, it should help to have working knowledge about the basic structure
of a spectrum generator, set out in \secref{sec:SpecGenStruct}.
In what follows,
it shall be demonstrated that
the clean object structure serves as firm guidance on the job.

@@

Consider a physics scenario which is best described
by a tower of effective theories.
Within the framework of \fs, the C++ object structure is
a faithful reflection of
this physicist's view on the given problem.
This point might be illustrated
using a well-known configuration in which
the higher-energy theory is the MSSMRHN discussed above which
gives rise to the MSSM as the lower-energy effective theory.
The relevant objects are sketched
in \figref{fig:class structure for tower}.
\begin{figure}
  \centering
\begin{tikzpicture}
  \draw (0,-.5) coordinate(bottom)
        (0,-.4)
     -- (0,0)  coordinate(MZ) node[left] {$M_Z$}
     -- (0,.7) coordinate(MS) node[left] {$M_S$}
     -- (0,2)  coordinate(Mv) node[left] {$M_\nu$}
     -- (0,3)  coordinate(MX) node[left] {$M_X$}
     -- (0,3.4)
        (0,3.5) coordinate(top);

@@

  \path (bottom) +(3.8,0) coordinate(SE) (top) +(-6.8,0) coordinate(NW);
  \draw  (NW) +(0,.5) rectangle (SE)
         (NW) -- +(10.6,0);
  \path (top) node[above] {\code{RGFlow}};
\end{tikzpicture}
  \caption{Schematic object structure in the C++ code for
    the tower scenario.}
  \label{fig:class structure for tower}
\end{figure}
The \code{MSSMRHN} object is in effect from the $M_X$ scale down to
the $M_\nu$ scale at which the right-handed neutrinos are decoupled.
Below this scale, the \code{MSSM} object takes over.
On the left of the vertical axis,
the boundary condition objects acting on either model are displayed,
together with the matching object connecting the two theories.
Note that each of the boundary condition and matching objects
maintains and updates its own scale over iterations.
An arrow in the figure depicts
the association of a constraint with its scale.
All these components are plugged into the \code{RGFlow} object
which then solves the problem.

Since the current implementation of \fs does not (yet)
produce a spectrum generator spanning multiple models,
one needs to write the matching object
% in \figref{fig:class structure for tower}
as well as gluing codes by hand to build such a program.
One can author nearly all the components of a multi-model spectrum generator
by making a straightforward extension
to each corresponding single-model counterpart
for one of the models forming the tower.
Obviously, this tip does not work for the matching class,
which has no analogue in a single-model case.

As the target spectrum generator depends on two models,
one should first build these prerequisites by:
\begin{numlstlisting}
$ ./createmodel --name=MSSM
$ ./createmodel --name=MSSMRHN
$ ./configure --with-models=MSSM,MSSMRHN
$ make
\end{numlstlisting}
As a by-product, line 3
also creates a \code{Makefile} in \code{examples/tower/}.
One could best see the overall code structure of the application in this file:
\begin{numlstlisting}
CPPFLAGS  := -I. $(INCCONFIG) $(INCFLEXI) $(INCLEGACY) $(INCSLHAEA) \
             $(INCMSSM) $(INCMSSMRHN)

@@

int main(int argc, char* argv[])
{
  // define objects;
  QedQcd oneset;
  MSSM_input_parameters    input_1;
  MSSMRHN_input_parameters input_2;
  // fill in input_1 and input_2;
  oneset.toMz(); // run SM fermion masses to MZ
  typedef Two_scale algorithm_type;
  MSSM_MSSMRHN_spectrum_generator<algorithm_type> spectrum_generator;
  // set up spectrum_generator;
  spectrum_generator.run(oneset, input_1, input_2);
  // extract outcome from models;
}
\end{numlstlisting}
where a line in the form \code{// ...;} shall be understood
to be a pseudo-code.
Given two models,
one declares two sets of input parameters,
\code{input_1} and \code{input_2}, in lines 7--8.
The suffixes \code{_1} and \code{_2} are not necessary but intended to remind
the programmer of to which model the object is relevant.

The \code{main} function then delegates
the \code{MSSM_MSSMRHN_spectrum_generator} object
to set up and drive the \code{RGFlow} object in
\figref{fig:class structure for tower}.
This task is started by the call in line 14 to
the following member function defined in
\code{MSSM_MSSMRHN_spectrum_generator.hpp}:

>>>>

The large variety of supersymmetric models and potential
investigations makes it likely that the user wants to modify the
generated spectrum generator source code or reuse components in
further programs.  \fs offers two levels to influence the code: (i) On
the \mathematica model file level the model itself or GUT/weak scale
boundary conditions as well as input and output parameters can be
controlled (see \secref{sec:quick-start-alternative-models} and
\secref{sec:adapting-model-files} for examples).  (ii) In particular
\fs uses C++ object orientation features to modularize the source code
so that it is sharply divided into building blocks performing
distinct duties.
This modular architecture makes it easy
for the user to modify, reuse, replace or
extend the individual components (see \secref{sec:adapting-cpp-code}
for examples).  An important application of this concept are the
boundary conditions, for which the C++ level offers a wider range of
possibilities.  The boundary conditions solver provides a plugin
mechanism via a common \code{Constraint} interface, which allows a
user to exchange or add boundary conditions at any scale.
To realized this, all (derived) constraint objects are intentionally
kept outside the solver.  Despite being independent of one another,
they can fit together with the aid of class inheritance.  An
elaborate example of a tower of effective field theories and multiple
matching scales is presented in \secref{sec:tower construction}.
Alternatively, the modular structure makes it straightforward to take
\fs generated code for e.g.\ RGEs or self-energies and reuse it in an
existing code for some other purpose.  Conversely, it is also easy to
include code from elsewhere into the spectrum generator.  For an
example see \secref{sec:integrating-custom-built}.

@@

By definition, research is an endeavor to find something new.
Therefore, it can often be the case that
a spectrum generator right out of the box is not enough.
\fs attempts to offer a clean interface through which
one can exploit its facilities
while undergoing a minimal amount of frustration,
when he or she programs for a wide variety of studies.
We provide two basic levels for the user to create a custom spectrum
generator: (i) The \mathematica level, where he or she writes or
adapts a model file and (ii) the C++ level, where the generated
classes can be extended, recombined or replaced by self-made modules.
In what follows, adaptions on these two levels shall be demonstrated
by presenting a few use cases at differing degrees of complexity.

@@

DEFINITION[EWSB][DiracSpinors] = {
  Fu  -> {FUL, conj[FUR]},
  Fv  -> {FV , conj[FV] },
  Chi -> {L0 , conj[L0] },
  Cha -> {Lm , conj[Lp] },
  ...
};
\end{numlstlisting}
With respect to the MSSM file, the newly added lines are
6, 12, 15, and 22.
Notice that the (left- and right-handed) neutrino mixing
in line 15 resembles the neutralino mixing in line 14.
Due to the Majorana mass term in the superpotential,
the six neutrino mass eigenstates are described in terms of
Majorana spinors like the neutralinos.

@@

WeylFermionAndIndermediate = {
  {v,   { Description -> "Right Neutrino Superfield" }},
  {FV,  { Description -> "Neutrino-Masseigenstate"}},
  {FvL, { Description -> "Left Neutrino"}},
  {FvR, { Description -> "Right Neutrino"}},
  ...
};
\end{numlstlisting}
In line 12, one finds PDG codes beginning with \code{99}.
Such numbers are available for a program author's private use
\cite{Beringer:1900zz}.
The new parameters in the superpotential
and the soft supersymmetry breaking sector
are to be described in \code{parameters.m}:
\begin{numlstlisting}
ParameterDefinitions = {
  {UV,    { Description -> "Neutrino-Mixing-Matrix"}},
  {Yv,    { Description -> "Neutrino-Yukawa-Coupling" }},
  {T[Yv], { Description -> "Trilinear-Neutrino-Coupling"}},
  {Mv,    { LaTeX -> "M_v", OutputName -> Mv, LesHouches -> Mv}},
  {B[Mv], { LaTeX -> "B_v", OutputName -> BMv, LesHouches -> BMv}},
  {mv2,   { Description -> "Softbreaking right Sneutrino Mass"}},
  ...
};
\end{numlstlisting}
For further details on how to write model files for \sarah, we refer
to its manual \cite{Staub:2008uz,Staub:2013tta}.

@@

There are problems which one cannot solve only by
editing \mathematica model files.
To unlock the full potential of \fs,
it is an advantage not to avoid programming at the C++ level.
For this, it should help to have working knowledge about the basic structure
of a spectrum generator, set out in \secref{sec:SpecGenStruct}.
In what follows,
it shall be demonstrated that
the clean class structure serves as firm guidance on the job.

@@

Consider a physics scenario which is best described
by a tower of effective theories.
Within the framework of \fs, the C++ class structure is
a faithful reflection of
this physicist's view on the given problem.
Here we illustrate this point
using a well-known configuration in which
the higher-energy theory is the MSSMRHN discussed above which
gives rise to the MSSM as the lower-energy effective theory.
The relevant classes are sketched
in \figref{fig:class structure for tower}.
\begin{figure}
  \centering
\begin{tikzpicture}
  \draw (0,-.5) coordinate(bottom)
        (0,-.4)
     -- (0,0)  coordinate(MZ) node[left] {$M_Z$}
     -- (0,.7) coordinate(MS) node[left] {$M_S$}
     -- (0,2)  coordinate(Mv) node[left] {$M_\nu$}
     -- (0,3)  coordinate(MX) node[left] {$M_X$}
     -- (0,3.4)
        (0,3.5) coordinate(top);

@@

  \path (bottom) +(3.8,0) coordinate(SE) (top) +(-6.8,0) coordinate(NW);
  \draw  (NW) +(0,.5) rectangle (SE)
         (NW) -- +(10.6,0);
  \path (top) node[above] {\code{RGFlow}};
\end{tikzpicture}
  \caption{Schematic class structure in the C++ code for
    the tower scenario.}
  \label{fig:class structure for tower}
\end{figure}
The \code{MSSMRHN} object is in effect from the $M_X$ scale down to
the $M_\nu$ scale at which the right-handed neutrinos are decoupled.
Below this scale, the \code{MSSM} object takes over.
On the left of the vertical axis,
the boundary condition objects acting on either model are displayed,
together with the matching object connecting the two theories.
Note that each of the boundary condition and matching objects
maintains and updates its own scale over iterations.
An arrow in the figure depicts
the association of a constraint with its scale.
All these components are plugged into the \code{RGFlow} object
which then solves the problem.

Since the current implementation of \fs does not (yet) automatically
produce a spectrum generator spanning multiple models,
one needs to write the matching class
% in \figref{fig:class structure for tower}
as well as gluing codes by hand to build such a program.
One can author nearly all the components of a multi-model spectrum generator
by making a straightforward extension
to each corresponding single-model counterpart
for one of the models forming the tower.
The matching class, however, needs to be set-up by hand, because it
has no analogue in a single-model case.

As the target spectrum generator depends on two models,
one should first build these prerequisites by:
\begin{numlstlisting}
$ ./createmodel --name=MSSM
$ ./createmodel --name=MSSMRHN
$ ./configure --with-models=MSSM,MSSMRHN
$ make
\end{numlstlisting}
As a by-product, line 3
also creates a \code{Makefile} in \code{examples/tower/}.
One can best see the overall code structure of the application in this file:
\begin{numlstlisting}
CPPFLAGS  := -I. $(INCCONFIG) $(INCFLEXI) $(INCLEGACY) $(INCSLHAEA) \
             $(INCMSSM) $(INCMSSMRHN)

@@

int main(int argc, char* argv[])
{
  // define objects;
  QedQcd oneset;
  MSSM_input_parameters    input_1;
  MSSMRHN_input_parameters input_2;
  // fill in input_1 and input_2;
  oneset.toMz(); // run SM fermion masses to MZ
  typedef Two_scale algorithm_type;
  MSSM_MSSMRHN_spectrum_generator<algorithm_type> spectrum_generator;
  // set up spectrum_generator;
  spectrum_generator.run(oneset, input_1, input_2);
  // extract outcome from models;
}
\end{numlstlisting}
where a line in the form \code{// ...;} shall be understood
to be a pseudo-code.
Given two models,
one declares two sets of input parameters,
\code{input_1} and \code{input_2}, in lines 7--8.

The crucial point is the definition of the
\code{MSSM_MSSMRHN_spectrum_generator} object in line 12, which
creates and drives the \code{RGFlow} object in \figref{fig:class
  structure for tower}.  This task is started by calling the
\code{run()} member function in line 14.  It is defined in
\code{MSSM_MSSMRHN_spectrum_generator.hpp} and reads:
